Chatbots and messaging platforms in the classroom: an analysis from the teacher's perspective

### 2021-06-27 edubots_surveys.Rnw
Reviewing State of the Art

Given the potential advantages that chatbots could bring to the classroom, as a previous step, we created two surveys for bachelor and master degree students at the University of Granada (Spain), which were answered by more than 250 students. The obtained results were analysed with the object of identifying students' actual needs, in \cite{MoraChatbos2021}. The survey asked about their preferences on tools/applications for chatting and messaging, how do they use them in the educational context, who do they like to be with in the class messaging groups, and what do they expect bots and chatbots to do to aid them during their learning process. We concluded that students prefer to use known whatever messaging application, such as Telegram or WhatsApp, they are already using. They also expected chatbots to help them with their scheduling, to provide assignment grades and to search for class material. Only one of the applications mentioned, having them work with frequently asked questions, might actually require some understanding of natural language, although it can be formulated in such a way that is not needed.

@@

WhatsApp is, at least in most Western countries,
the most popular MIM platform used by educators to give assignments'
feedback to students, support course discussions and provide learning
resources in informal learning settings \cite{panah2020study}.  The
use of WhatsApp in higher education could enhance social presence
\cite{tang2017mobile} and foster trust relationships between educators
and students embedded in the social learning process
\cite{gachago2015crossing}.

Nevertheless, there are challenges when using MIM tools that occur due
to the blurring of boundaries between academic and private life. This
can lead to technostress \cite{gachago2015crossing}, difficulty in
managing responsibilities, especially among mature students, and lack of
privacy \cite{tang2017mobile}.  Students' dropout of the MIM groups,
as they can leave groups at any time, can hinder their learning and
undermine educators' efforts \cite{mwakapina2016whatsapp}.  In
addition, there is a need to set rules and norms for these MIM groups
in order to maintain the safety of these online communities for
students \cite{abbas2021onlinechat} without affecting students'
ownership and control that is vital to advance their learning
\cite{gachago2015crossing}. This is why examining the role of MIM in higher education is still a challenge, and why the attitudes of the teaching community towards them has to be examined, as we do in this paper.

% The Use of Chatbots in Higher Education
The use of MIM, although possibly valuable by itself, can be enhanced via the use of chatbots, which, being conversational agents, usually well where synchronous conversations take place. 
The use of conversational agents or chatbots in higher education is
still at its infancy \cite{yang2019opportunities}. Nevertheless,
recent studies examining their positive impact on students' academic
performance \cite{perez2020rediscovering} and engagement
\cite{differ.chat,abbas2021onlinechat} have led to a growing interest
in using this technology in the (possibly virtual) classroom. Indeed, the use of chatbots in
collecting course feedback from students in higher education improved
students' response quality and boosted their enjoyment levels
\cite{abbas2021Surveys}.  According to \cite{Roblyer2010}, using
either tools such as mobile devices or teaching strategies based on
gamification \cite{Yildirim2017} can improve student motivation. In
this sense, \cite{Pimmer2019} carried out a series of experiments with
technological tools whose results showed positive perception and
acceptance of the use of technology in teaching and learning.

Several higher education chatbots' evaluation studies were
undertaken. For instance, an evaluation review study presented by
\cite{Smutny2020} examined 47 educational chatbots implemented in
Facebook Messenger with the focus to identify characteristics and
quality metric such as language, subject matter and platform. Another
study aimed to categorise educational chatbots according to their
purpose into service-oriented and teaching-oriented
\cite{perez2020rediscovering}. The first category includes those that
provide service support such as the chatbot Ask Holly \cite{meetholly}
and Dina \cite{santoso2018dinus}, both chatbots respond to students'
questions about enrollment and registration. Ask
L.U. \cite{LancasterUniversity_2019} answers students' frequently
asked questions about timetables, grades, tutors and societies. LISA
\cite{dibitonto2018chatbot} and Differ \cite{differ.chat} facilitate
breaking the ice between new students by introducing them to each
other.  \cite{ranoliya2017chatbot} proposed a generic chatbot for
university students that is able to answer their frequently asked
questions; for some time, the University of Granada had ELVIRA as a
chatbot in their main page \cite{MOREO20129}. Besides being able to
answer pre-established (frequent) questions, it did so from the main
University of Granada using an inset character, voiced by a real
person, who lip-synched the answers. The emphasis in this case,
however, was on the authenticity of the synchronization more than in
making the update of questions and answers simpler, or more
interactive, using actual messaging platforms. In most cases, using
the inset search engine was able to give you more up-to-date and
accurate answers.

On the other hand, teaching-oriented chatbots are more sophisticated
as they set personalised learning outcomes and monitor learning
progress. For instance, \cite{fernoagua2018intelligent} reported on
''eduAssistant'' a virtual teaching assistant chatbot developed on the
Telegram messaging platform. In this study, the Telegram platform was
chosen because it is easy to use, students are familiar with its
features and it enables students to exchange messages in different
formats (text, audio and video) \cite{fernoagua2018intelligent}. In
addition, Telegram could operate on all devices and operating
systems. The ''eduAssistant'' chatbot acts as an automatic agent
between the teacher-content-student, facilitating real-time feedback
loops and providing a personalised learning experience relevant to
students' skills and knowledge. Using this chatbot, educators can
create interactive instances in their lectures where they pose
questions to their students and the chatbot assists students who need
further help by giving those more hints and reporting to their
educator's dashboard \cite{fernoagua2018intelligent}. This can help
educators locate those students that need their attention and send
them more educational resources relevant to their academic
attainment.

Despite its (arguably) successful implementation in different higher education institutions, over all in the fringes of educational activity, its implementation or deployment is not trivial.
Some authors \cite{sjostrom2018designing} have proposed a conceptual
architecture for the adoption teaching-oriented chatbots in higher education.  This
conceptual architecture is based on a systematic literature review of
previous studies examining the design of chatbots in higher education
and a content analysis of student emails and discussion forum posts of
four instances of a Java programming course.  The study outlined
several design considerations; among them, the authors emphasised the
importance of developing chatbots in platforms that students and
educators are familiar with and can easily access (i.e. Facebook
Messenger) which confirms with \cite{hobert2019you} and
\cite{fernoagua2018intelligent} studies. In addition,
\cite{sjostrom2018designing} argued that a conceptualision of
learners' questions could aid designers in integrating the appropriate
types of questions that the chatbots should support for different
courses.  \cite{coronado2018cognitive} proposed agents that store
learning materials to be provided on demand to
students. \cite{crockett2017predicting} reported on tutoring systems,
which can perform initial assessments of students' understanding and
provide learning material that would advance their understanding to
the next level.

Regarding the factors for the adoption of chatbots in higher
education, many studies have focused on the evaluation of technology
acceptance and usability \cite{Roblyer2010,Pimmer2019}. However,
higher education is a special domain where according to
\cite{hobert2019you}, specific pedagogical factors such as learning
success or increased motivation are more important.  Therefore, to
develop efficient chatbots for higher education, all stakeholders'
(i.e. educators, students, institutions, etc) expectations needs to be
carefully collected and taken into consideration
\cite{sjostrom2018designing,tsivitanidouusers}.

@@

Our intention from the beginning was to try and gather a sufficient amount of responses in all sectors, as well as cover different sectors of higher education. This is why we created a Google Form, in Spanish, which was intended to be answered in a very short amount of time. Initially, our intention was to mirror the questions used in \cite{MoraChatbos2021}; however, an initial exploration of results led us to create a second survey.
% Questions should probably be included here.

>>>>

Given the potential advantages that chatbots could bring to the classroom, as a previous step, we created two surveys for bachelor and master degree students at the University of Granada (Spain), which were answered by more than 250 students. The obtained results were analysed with the object of identifying students' actual needs, in \cite{MoraChatbots2021}. The survey asked about their preferences on tools/applications for chatting and messaging, how do they use them in the educational context, who do they like to be with in the class messaging groups, and what do they expect bots and chatbots to do to aid them during their learning process. We concluded that students prefer to use known whatever messaging application, such as Telegram or WhatsApp, they are already using. They also expected chatbots to help them with their scheduling, to provide assignment grades and to search for class material. Only one of the applications mentioned, having them work with frequently asked questions, might actually require some understanding of natural language, although it can be formulated in such a way that is not needed.

@@

WhatsApp is, at least in most Western countries,
the most popular MIM platform used by educators to give assignments'
feedback to students, support course discussions, and provide learning
resources in informal learning settings \cite{panah2020study}.  Moreover, the
use of WhatsApp in higher education could enhance social presence
\cite{tang2017mobile} and foster trust relationships between educators
and students embedded in the social learning process
\cite{gachago2015crossing}.

Nevertheless, there are challenges when using MIM tools that occur due
to the blurring of boundaries between academic and private life. This
can lead to technostress \cite{gachago2015crossing}, difficulty in
managing responsibilities, especially among mature students, and lack of
privacy \cite{tang2017mobile}.  Students' dropout of the MIM groups,
as they can leave groups at any time, can hinder their learning and
undermine educators' efforts \cite{mwakapina2016whatsapp}.  In
addition, there is a need to set rules and norms for these MIM groups
in order to maintain the safety of these online communities for
students \cite{abbas2021onlinechat}. However, these rules should not affect students'
ownership and control, since it is vital to advance in their learning
\cite{gachago2015crossing} precess. 
This is why examining the role of MIM in higher education is still a challenge, and why the attitudes of the teaching community towards them have to be examined, as we do in this paper.

% The Use of Chatbots in Higher Education
The use of MIM, although possibly valuable by itself, can be enhanced via the use of chatbots, which, being conversational agents, usually well where synchronous conversations take place. 
% Antonio - I cannot understand this sentence... what's the point of the expression 'usually well'?
The use of conversational agents (chatbots) in higher education is
still at its infancy \cite{yang2019opportunities}. Nevertheless,
recent studies examining their positive impact on students' academic
performance \cite{perez2020rediscovering} and engagement
\cite{differ.chat,abbas2021onlinechat} have led to a growing interest
in using this technology in the (possibly virtual) classroom. Indeed, the use of chatbots in
collecting course feedback from students in higher education improved
students' response quality and boosted their enjoyment levels
\cite{abbas2021Surveys}.  According to \cite{Roblyer2010}, using
either tools such as mobile devices or teaching strategies based on
gamification \cite{Yildirim2017} can improve student motivation. In
this sense, the authors in \cite{Pimmer2019} carried out a series of experiments with
technological tools whose results showed positive perception and
acceptance of the use of technology in teaching and learning.

Several higher education chatbots' evaluation studies have been 
undertaken. For instance, a recent evaluation review study presented by Smutny and Schreiberova 
\cite{Smutny2020} examined 47 educational chatbots implemented in
Facebook Messenger with the focus to identify characteristics and
quality metrics such as language, subject matter and platform. Another
study aimed to categorise educational chatbots according to their
purpose into service-oriented and teaching-oriented
\cite{perez2020rediscovering}. The first category includes those that
provide service support such as the chatbot Ask Holly \cite{meetholly}
and Dina \cite{santoso2018dinus}; both chatbots respond to students'
questions about enrollment and registration. Ask (Lancaster University)
 \cite{LancasterUniversity_2019} answers students' frequently
asked questions about timetables, grades, tutors and societies. LISA
\cite{dibitonto2018chatbot} and Differ \cite{differ.chat} facilitate
breaking the ice between new students by introducing them to each
other. Ranoliya et al. \cite{ranoliya2017chatbot} proposed a generic chatbot for
university students that is able to answer to their frequently asked
questions. 
For some time, the University of Granada had ELVIRA as a
chatbot in its main web page \cite{MOREO20129}. Besides being able to
answer pre-established (frequent) questions, it did so from the website of the
University of Granada using an inset character, voiced by a real
person, who lip-synched the answers. The emphasis in this case,
however, was on the authenticity of the synchronization more than in
making the update of questions and answers simpler, or more
interactive, using actual messaging platforms. 
% Antonio - the last part of the sentence is not clear. Do you refer to all the aforementioned chatbots?
% Just to Elvira? What messaging platforms do you mean?
In most cases, using the inset search engine was able to give you more up-to-date and
accurate answers.
% Antonio - again, are you speaking about Elvira? 
% If we say something like the last sentence, we are actually saying that chatbots are not useful... is this what we want to remark?

On the other hand, teaching-oriented chatbots are more sophisticated,
as they set personalised learning outcomes and monitor learning
progress. For instance, the work \cite{fernoagua2018intelligent} reported on
''eduAssistant'', a virtual teaching assistant chatbot developed inside the
Telegram messaging platform. In this study, Telegram platform was
chosen because it is easy to use, students are familiar with its
features, and it enables users to exchange messages in different
formats (text, audio and video) \cite{fernoagua2018intelligent}. In
addition, Telegram run on all devices and operating
systems. ''eduAssistant'' chatbot acts as an automatic agent
in teacher-content-student, facilitating real-time feedback
loops and providing a personalised learning experience relevant to the
students' acquired skills and knowledge. Using this chatbot, educators can
create interactive instances in their lectures where they pose
questions to their students and the chatbot assists those who need
further help by giving them more hints and reporting it to their
educator's dashboard \cite{fernoagua2018intelligent}. This can help 
educators to locate those students that need their attention and send to 
them more educational resources relevant to their academic attainment.

Despite its (arguably) successful implementation in different higher education institutions, 
over all in the fringes of educational activity, its implementation or deployment is not trivial.
Authors in \cite{sjostrom2018designing} have proposed a conceptual
architecture for the adoption of teaching-oriented chatbots in higher education.  This
conceptual architecture is based on a systematic literature review of
previous studies, examining the design of chatbots in higher education, as well as a content 
analysis of student emails and discussion forum posts of
four instances of a Java programming course.  The study outlined
several design considerations; among them, the authors emphasised the
importance of developing chatbots in platforms that students and
educators are familiar with and can easily access (i.e. Facebook
Messenger) which was confirmed with \cite{hobert2019you} and
\cite{fernoagua2018intelligent} studies. In addition, Sj{\"o}str{\"o}m et al. \cite{sjostrom2018designing} 
argued that a conceptualision of learners' questions could aid designers in integrating the appropriate
types of questions that the chatbots should support for different
courses.  Coronado et al. \cite{coronado2018cognitive} proposed agents that store
learning materials to be provided on demand to students, whereas the authors in
\cite{crockett2017predicting} reported on tutoring systems,
which can perform initial assessments of students' understanding and
provide learning material that would advance their understanding to
the next level.

Regarding the factors for the adoption of chatbots in higher
education, many studies have focused on the evaluation of technology
acceptance and usability \cite{Roblyer2010,Pimmer2019}. However,
higher education is a special domain where according to Hobert \cite{hobert2019you}, 
specific pedagogical factors such as learning success or increased motivation are more important.  
Therefore, to develop efficient chatbots for higher education, all stakeholders'
(i.e. educators, students, institutions, etc) expectations needs to be
carefully collected and taken into consideration
\cite{sjostrom2018designing,tsivitanidouusers}.

In this line, our previous work \cite{MoraChatbots2021} aimed at analysing the expectations of students on this regard, whereas the present study is focused on the other key actor in this challenge, i.e. educators.

@@

Our intention from the beginning was to try and gather a sufficient amount of responses in all sectors, as well as cover different sectors of higher education. This is why we created a Google Form, in Spanish, which was intended to be answered in a very short amount of time. Initially, our intention was to mirror the questions used in \cite{MoraChatbots2021}; however, an initial exploration of results led us to create a second survey.
% Questions should probably be included here.
